Laplacian envelope

### 2020-06-18 Notes/laplace.tex
details re shortening chain

In analogy with the case $s=0$, we define
%
\begin{equation}\label{eq:lareacoeffs}
  \begin{aligned}
    c_k(s) &= \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)}, \\
    a_j(s) &= \sum_{jk} \eq_j \enc_{jk} (\etw_{j}(s) - \etw_{k}(s)), \\
    \thb_i(s) &= \sum_j \fptb_{ij}(0)\ a_j(s), \\
  \end{aligned}
    \implies
    \hat{A}(s) = \sum_k c_k(s) \eq_k \w_k
      = \sum_i a_i(s).
\end{equation}
%

@@

To write down the Hessian, we will first define some notation.
We define the matrix $\mathbf{X}(s) = \fund(0) \enc \fund(s)$.
Given a four-index array $H_{ijkl}$, we define $H\trans_{ijkl} = H_{klij}$.
Then we denote
%
\begin{equation}\label{eq:hessparts}
\begin{aligned}
  R_{ijkl} &=
    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\thb_k(s) - \thb_l(s)} \\
    &&&+ \eq_i \brc{\mathbf{X}_{ik}(s) - \mathbf{X}_{jk}(s)} \brc{\etw_k(s) - \etw_l(s)} \\
    &&&+ \eq_i c_i(s) \brc{\fund_{ik}(s) - \fund_{jk}(s)} \brc{\etw_k(s) - \etw_l(s)},  \\
  S_{ijkl} &=
    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\etw_k(s) - \etw_l(s)} \\
    &&&+ \eq_k \brc{\fund_{ki}(s) - \fund_{li}(s)} \brc{\etw_i(s) - \etw_j(s)}.
\end{aligned}
\end{equation}
%
Then we can write the Hessian as
%
\begin{equation}\label{eq:hessian}
\begin{aligned}
  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\pot_{kl}} &=
     (f\pot)^2   [R + R\trans + S + S\trans]_{ijkl}, \\
  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\dep_{kl}} &=
     f\pot f\dep [R + R\trans - S + S\trans]_{ijkl}, \\
  \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\pot_{kl}} &=
     f\pot f\dep [R + R\trans + S - S\trans]_{ijkl}, \\
  \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\dep_{kl}} &=
     (f\dep)^2   [R + R\trans - S - S\trans]_{ijkl}, \\
\end{aligned}
\end{equation}
%
This is useful in numerical maximisation.

@@

In reality, $M$ can only take even integer values.
However, one can find a heuristic approximation to the envelope by maximising \wrt $M$.
We find that
%
\begin{multline}\label{eq:unienv}
  s = 2\sinh^2\frac{y_*}{M_*},
  \qquad
  \hat{A}_*(s) = \prn{ \frac{ 4\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } } \frac{\log\alpha(s)}{s},
  \\ \text{where}\quad
  y_* = \tanh \frac{y_*}{2} \cosh y_* \approx 1.50553.
  \qquad
\end{multline}
%
This is only valid for
%
\begin{equation}\label{eq:univalid}
  2 \leq M_* \leq M,
  \qquad\implies\qquad
  2\sinh^2 \frac{y_*}{M} \leq s \leq 2\sinh^2 \frac{y_*}{2}.
\end{equation}
%

When $M \gg 2$, the validity region is approximately
%
\begin{equation}\label{eq:univalidapprox}
  \frac{4.53327}{M^2} \lesssim s \lesssim 1.36423.
\end{equation}
%
When $s \sim \CO(M^{-2}) \ll 1$, we have $\alpha \approx 1+\sqrt{2s}$ and the heuristic envelope is approximately
%
\begin{equation}\label{eq:unienvapprox}
  \hat{A}_*(s) \approx \frac{ 1.08406 }{ \sqrt{s} }.
\end{equation}
%

@@

The following elements of $\enc$ and $\eq$ are differ from \eqref{eq:unienc}:
The following elements of $\enc$ and $\eq$ are differ from \eqref{eq:unienc}:
%
\begin{equation}\label{eq:shortenenc}
  \begin{aligned}
  \enc_{M-1M} = -\enc_{21} &= \frac{\varepsilon}{2},
  &\qquad
  \eq_1 = \eq_M &= \frac{\varepsilon}{M-2(1-\varepsilon)},
  \\ &&
  \eq_i &=  \frac{1}{M-2(1-\varepsilon)}
  \quad\text{for}\quad 1<i<M.
  \end{aligned}
\end{equation}
%
This means that the Laplace transform is given by
%
\begin{equation}\label{eq:shortenareaeta}
  \hat{A}(s) = \frac{\varepsilon(\etw_1(s)-\etw_M(s)) + (1-\varepsilon)(\etw_2(s)-\etw_{M-2}(s))}{M-2(1-\varepsilon)}
       = \frac{2\varepsilon\, \etw_1(s) + 2(1-\varepsilon)\etw_2(s)}{M-2(1-\varepsilon)}.
\end{equation}
%

The $\etw_i(s)$ take the same general form \eqref{eq:detagensol} except for $i=1$, and with different coefficients due to the different boundary conditions
%
\begin{equation}\label{eq:detabndshorten}
\begin{aligned}
  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_3 + \varepsilon\,\delta\eta_1}{2} - \prn{s+\frac{1+\varepsilon}{2}}\delta\eta_2 &= 0, \\
  \frac{\delta\eta_{\frac{M}{2}-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{\frac{M}{2}} &= \frac{1}{s}.
\end{aligned}
\end{equation}
%
These are solved by
%
%
\begin{equation}\label{eq:shortensol}
\begin{aligned}
  \delta\eta_i &= B\alpha^{i-2} + C\alpha^{2-i} \qquad
        \text{for} \quad 2 \leq i \leq \frac{M}{2}, \\
  \delta\eta_1 &= \frac{2\alpha^{\frac{M}{2}}}
         {[\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
  B &= - \frac{2\alpha^{\frac{M}{2}}(\alpha^2-\alpha+1 + \varepsilon(\alpha-1))}
         {(1+\alpha)[\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
  C &= - \frac{2\alpha^{\frac{M-2}{2}}(\alpha^2-\alpha+1 - \varepsilon\alpha(\alpha-1))}
         {(1+\alpha)[\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
\end{aligned}
\end{equation}
%
Which gives the following Laplace transform
%
\begin{equation}\label{eq:shortenlaplace}
  \hat{A}(s) = \frac{2 [\varepsilon(\alpha^{\frac{M}{2}}-1)^2
          + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2] }
          { [M-2(1-\varepsilon)]
            [\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s }.
\end{equation}
%
This interpolates between \eqref{eq:unilaplace} at $\varepsilon=1$ and the same expression for $M-2$ at $\varepsilon=0$.

We can maximise this \wrt $\varepsilon$.
This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
We can find the boundary of the region of applicability by asking at what value of $s$ is the optimal value of $\varepsilon=0$ or $1$?
%
\begin{equation}\label{eq:shortenvalid}
\begin{aligned}
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=1} \propto &\,
    M(\alpha^2-\alpha+1)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^M-1) - (\alpha^{\frac{M}{2}}-1)^2(\alpha^{M}+1)
      = 0,\\
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=0} \propto &\,
    (M-2)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^M-1) - (\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2(\alpha^{M-2}+1)
      = 0,\\
    s = &\, \frac{(\alpha-1)^2}{2\alpha},
\end{aligned}
\end{equation}
%
with the $\varepsilon=1$ case providing the lower boundary and $\varepsilon=0$ providing the upper boundary.
If these were really the models that maximised $A(s)$, the region of validity for $M$ would not overlap with that for $M-2$.
In this case, however, there is an overlap.
This is related to the fact that \autoref{fig:envvid} and the video shows that the optimal models do not shorten by one state at a time, but several outgoing transition probabilities are reduced at once.

@@

We will phrase it in terms of $\snrb(\tau)$, reintroducing $r$:
%
\begin{equation}\label{eq:envunits}
  \snrb(\tau) = \frac{ \sqrt{N} \hat{A}\prn{\frac{1}{r\tau}} }{ r\tau },
  \qquad
  \alpha(\tau) = \frac{ 1 + r\tau +\sqrt{1+2r\tau} }{ r\tau }.
\end{equation}
%
Then, combining \eqref{eq:binaryenv}, \eqref{eq:unienv} and \eqref{eq:stickylaplace}, we find
%
\begin{equation}\label{eq:henv}
  \snrb(\tau) = \left\{
\begin{aligned}
  & \frac{\sqrt{N}}{1+r\tau}, &
  & r\tau \leq \frac{1}{2\sinh^2\frac{y_*}{2}} , \\
  & \sqrt{N}\prn{ \frac{ 4\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } } \log\alpha, &
  \frac{1}{2\sinh^2 \frac{y_*}{2}} \leq & r\tau \leq \frac{1}{2\sinh^2 \frac{y_*}{M}} , \\
  & \max_\varepsilon \brc{\frac{2\sqrt{N}\varepsilon}{2+(M-2)\varepsilon} \,
      \frac{ (\alpha^{\frac{M}{2}}-1)^2 - (1-\varepsilon)\alpha(\alpha^{\frac{M-2}{2}}-1)^2 }{ (\alpha^M + 1) - (1-\varepsilon)\alpha(\alpha^{M-2}+1) }}, &
  \frac{1}{2\sinh^2 \frac{y_*}{M}} \leq & r\tau .
\end{aligned}
\right.
\end{equation}
%
For $M \gg 1$, and $r\tau$ much greater than the lower bound of each region, we can approximate this as
%
\begin{equation}\label{eq:henvapproc}
  \snrb(\tau) \approx \left\{
\begin{aligned}
  & \frac{\sqrt{N}}{1+r\tau}, &
  & r\tau \leq 0.73 , \\
  & \frac{ 1.1\sqrt{N}}{ \sqrt{r\tau} } , &
  0.73 \leq & r\tau \leq 0.22 M^2 , \\
  & \frac{ \sqrt{N} (M-1) }{ r\tau }, &\quad
  0.22 M^2 \leq &r\tau .
\end{aligned}
\right.
\end{equation}
%
This is plotted in \autoref{fig:heuristicenv}.

>>>>

In analogy with the case $s=0$, we define
%
\begin{equation}\label{eq:lareacoeffs}
  \begin{aligned}
    c_k(s) &= \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)}, \\
    a_j(s) &= \sum_{k} \enc_{jk} (\etw_{j}(s) - \etw_{k}(s)), \\
    \thb_i(s) &= \sum_j \fptb_{ij}(0)\, \eq_j a_j(s), \\
  \end{aligned}
    \implies
    \hat{A}(s) = \sum_k c_k(s) \eq_k \w_k
      = \sum_i  \eq_i a_i(s).
\end{equation}
%

@@

To write down the Hessian, we will first define some notation.
We define the matrix $\mathbf{X}(s) = \fund(0) \enc \fund(s)$.
Given a four-index array $H_{ijkl}$, we define $H\trans_{ijkl} = H_{klij}$.
Then we denote
%
\begin{equation}\label{eq:hessparts}
\begin{alignedat}{3}
    R_{ijkl} &= \
    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\thb_k(s) - \thb_l(s)} \\
    &&&+ \eq_i \brc{\mathbf{X}_{ik}(s) - \mathbf{X}_{jk}(s)} 
          \brc{\etw_k(s) - \etw_l(s)} \\
    &&&+ \eq_i c_i(s) \brc{\fund_{ik}(s) - \fund_{jk}(s)} 
          \brc{\etw_k(s) - \etw_l(s)},  \\
  S_{ijkl} &=
    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\etw_k(s) - \etw_l(s)} \\
    &&&+ \eq_k \brc{\fund_{ki}(s) - \fund_{li}(s)} \brc{\etw_i(s) - \etw_j(s)}.
\end{alignedat}
\end{equation}
%
Then we can write the Hessian as
%
\begin{equation}\label{eq:hessian}
\begin{aligned}
  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\pot_{kl}} &=
     (f\pot)^2   [R + R\trans + S + S\trans]_{ijkl}, \\
  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\dep_{kl}} &=
     f\pot f\dep [R + R\trans - S + S\trans]_{ijkl}, \\
  \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\pot_{kl}} &=
     f\pot f\dep [R + R\trans + S - S\trans]_{ijkl}, \\
  \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\dep_{kl}} &=
     (f\dep)^2   [R + R\trans - S - S\trans]_{ijkl}, \\
\end{aligned}
\end{equation}
%
This is useful in numerical maximisation.

@@

In reality, $M$ can only take even integer values.
However, one can find a heuristic approximation to the envelope by maximising \wrt $M$.
We find that
%
\begin{multline}\label{eq:unienv}
  \alpha = \exp\prn{\pm\frac{2y_*}{M_*}},
  \quad
  s = 2\sinh^2\frac{y_*}{M_*},
  \qquad
  \hat{A}_*(s) = \prn{\frac{ 2\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } } \frac{\abs{\log\alpha(s)}}{s},
  \\ \text{where}\quad
  y_* = \tanh \frac{y_*}{2} \cosh y_* \approx 1.50553.
  \qquad
\end{multline}
%
This is only valid for
%
\begin{equation}\label{eq:univalid}
  2 \leq M_* \leq M,
  \qquad\implies\qquad
  2\sinh^2 \frac{y_*}{M} \leq s \leq 2\sinh^2 \frac{y_*}{2}.
\end{equation}
%

When $M \gg 2$, the validity region is approximately
%
\begin{equation}\label{eq:univalidapprox}
  \frac{4.53327}{M^2} \lesssim s \lesssim 1.36423.
\end{equation}
%
When $s \sim \CO(M^{-2}) \ll 1$, we have $\alpha \approx 1 \pm \sqrt{2s}$ and the heuristic envelope is approximately
%
\begin{equation}\label{eq:unienvapprox}
  \hat{A}_*(s) \approx \frac{ 0.54203 }{ \sqrt{s} }.
\end{equation}
%

@@

The following elements of $\enc$ and $\eq$ differ from \eqref{eq:unienc}:
%
\begin{equation}\label{eq:shortenenc}
  \begin{aligned}
  \enc_{M-1M} = -\enc_{21} &= \frac{(1-\varepsilon)}{2},
  &\qquad
  \eq_1 = \eq_M &= \frac{(1-\varepsilon)}{M-2\varepsilon},
  \\ &&
  \eq_i &=  \frac{1}{M-2\varepsilon}
  \quad\text{for}\quad 1<i<M.
  \end{aligned}
\end{equation}
%
This means that the Laplace transform is given by
%
\begin{equation}\label{eq:shortenareaeta}
  \hat{A}(s) = \frac{(1-\varepsilon)(\etw_1(s)-\etw_M(s)) + \varepsilon(\etw_2(s)-\etw_{M-1}(s))}{M-2\varepsilon}
       = \frac{2(1-\varepsilon) \etw_1(s) + 2\varepsilon\,\etw_2(s)}{M-2\varepsilon}.
\end{equation}
%

The $\etw_i(s)$ take the same general form \eqref{eq:detagensol} except for $i=1$, and with different coefficients due to the different boundary conditions
%
\begin{equation}\label{eq:detabndshorten}
\begin{aligned}
  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_3 + (1-\varepsilon)\delta\eta_1}{2} - \prn{s+\frac{2-\varepsilon}{2}}\delta\eta_2 &= 0, \\
  \frac{\delta\eta_{\frac{M}{2}-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{\frac{M}{2}} &= \frac{1}{s}.
\end{aligned}
\end{equation}
%
These are solved by
%
%
\begin{equation}\label{eq:shortensol}
\begin{aligned}
  \delta\eta_i &= B\alpha^{i-2} + C\alpha^{2-i} \qquad
        \text{for} \quad 2 \leq i \leq \frac{M}{2}, \\
  \delta\eta_1 &= \frac{2\alpha^{\frac{M}{2}}}
         {[(1-\varepsilon)(\alpha^{M}+1) + \varepsilon(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
  B &= - \frac{2\alpha^{\frac{M}{2}}(\alpha^2-\alpha+1 + (1-\varepsilon)(\alpha-1))}
         {(1+\alpha)[(1-\varepsilon)(\alpha^{M}+1) + \varepsilon(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
  C &= - \frac{2\alpha^{\frac{M-2}{2}}(\alpha^2-\alpha+1 - (1-\varepsilon)\alpha(\alpha-1))}
         {(1+\alpha)[(1-\varepsilon)(\alpha^{M}+1) + \varepsilon(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
\end{aligned}
\end{equation}
%
Which gives the following Laplace transform
%
\begin{equation}\label{eq:shortenlaplace}
  \hat{A}(s) = \frac{2 [(1-\varepsilon)(\alpha^{\frac{M}{2}}-1)^2
          + \varepsilon(\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2] }
          { [M-2(1-\varepsilon)]
            [(1-\varepsilon)(\alpha^{M}+1) + \varepsilon(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s }.
\end{equation}
%
This interpolates between \eqref{eq:unilaplace} at $\varepsilon=0$ and the same expression for $M-2$ at $\varepsilon=1$.
This is easier to see if we define
%
\begin{equation}\label{eq:shortdefs}
\begin{aligned}
  F_M(\alpha) &= (\alpha^{\frac{M}{2}} - 1)^2, &
  \Delta F_M(\alpha) &= F_M(\alpha) - \zeta F_{M-2}(\alpha), &
  f_m(\alpha) = \frac{F_M(\alpha)}{\Delta F_M(\alpha)}, \\
  G_M(\alpha) &= \alpha^{M} + 1, &
  \Delta G_M(\alpha) &= G_M(\alpha) - \zeta G_{M-2}(\alpha), &
  g_m(\alpha) = \frac{G_M(\alpha)}{\Delta G_M(\alpha)}, \\
  \text{where}\quad \zeta &= \alpha^2 - \alpha + 1, &
  m &= \frac{M}{2}. \\
\end{aligned}
\end{equation}
%
Then the uniform model \eqref{eq:unilaplace} has \( \hat{A}(s) = \frac{F_m(\alpha)}{s m G_m(\alpha)} \).
Then \cref{eq:shortenlaplace} can be be written as
%
\begin{equation}\label{eq:shortsimple}
  \hat{A}(s) =
    \frac{(1-\varepsilon)F_m(\alpha) + \varepsilon \zeta F_{m-1}(\alpha)}
         {s (m - \varepsilon) ((1-\varepsilon)G_m(\alpha) + \varepsilon \zeta G_{m-1}(\alpha))}
    % = \frac{F_m - \varepsilon \Delta F_{m}}
    %      {s (m - \varepsilon) (G_m - \varepsilon \Delta G_{m})}
    = \brk{\frac{\Delta F_{m}}{s \Delta G_{m}}}
    \frac{f_m - \varepsilon}{(m - \varepsilon) (g_m - \varepsilon)}.
\end{equation}
%

\begin{figure}[ht]
\begin{center}
\begin{myenuma}
  \item\label{fig:short_opt}\aligntop{%
    \includegraphics[width=0.45\linewidth]{shorten_opt.pdf}}
  \item\label{fig:short_ratio}\aligntop{%
    \includegraphics[width=0.45\linewidth]{shorten_ratio.pdf}}
\end{myenuma}
\caption{Signal-to-noise ratio for optimal shortened models, \cref{eq:shortenenc} optimised over \(\varepsilon\).
(\ref{fig:short_opt}): Solid grey line is \cref{eq:unienv}, the uniform model with \(M\) treated as a continuous parameter and optimised.
Dotted grey line is the approximate expression from \cref{eq:unienvapprox}.
The other curves are from \cref{eq:shortopt}, optimised shortened models for various values of \(M\).
(\ref{fig:short_ratio}): The ratios of the quantities plotted in (\ref{fig:short_opt}) to \cref{eq:unienv}.
The dotted orange line is the extension of the \(M=4\) line with \(\varepsilon\) held at 0, as discussed below \cref{eq:shortenvalid}.
\label{fig:shorten}}
\end{center}
\end{figure}

We can maximise this \wrt $\varepsilon$.
%
\begin{equation}\label{eq:shortopt}
  (\varepsilon - f_m)^2 = (m - f_m)(g_m - f_m),
  \qquad
  \hat{A}(s) = \brk{\frac{\Delta F_{m}}{s \Delta G_{m}}}
    \frac{1}{\brk{\sqrt{\abs{m - f_m}} - \sqrt{\abs{g_m - f_m}}}^2}.
\end{equation}
%
We can find the boundary of the region of applicability by asking at what value of $s$ is the optimal value of $\varepsilon=0$ or $1$?
%
\begin{equation}\label{eq:shortenvalid}
\begin{aligned}
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=0} \propto &\,
    M(\alpha^2-\alpha+1)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^{M-1}-1) - (\alpha^{\frac{M}{2}}-1)^2(\alpha^{M}+1)
      = 0,\\
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=1} \propto &\,
    (M-2)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^{M-1}-1) - (\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2(\alpha^{M-2}+1)
      = 0,\\
    \text{where}\quad
    s = &\, \frac{(\alpha-1)^2}{2\alpha},
\end{aligned}
\end{equation}
%
with the $\varepsilon=0$ case providing the lower boundary and $\varepsilon=1$ providing the upper boundary.
For lower \(s\), \(\varepsilon = 0\) is the optimum. 
For higher \(s\), \(\varepsilon = 1\) is the optimum, which is eqivalent to \(\varepsilon = 0\) with \(M-2\) states.
This is depicted by the dotted orange line in \cref{fig:shorten}\ref{fig:short_ratio}.

\(M=2\) is a special case:
%
\begin{equation*}
  \hat{A}(s) = \frac{(\alpha-1)^2}
                    {[2(\alpha^2-\alpha+1) - (1-\varepsilon) (\alpha-1)^2]s},
\end{equation*}
%
so \(\varepsilon=0\) is always the maximum.

If these were really the models that maximised $A(s)$, the region of validity for $M$ would not overlap with that for $M-2$.
In this case, however, there is an overlap as seen in \cref{fig:shorten}\ref{fig:short_ratio}.
The \(M=2\) region of validity overlaps everything else.
The other overlaps are less extreme, but they do exist.
This is related to the fact that \autoref{fig:envvid} and the video shows that the optimal models do not shorten by one state at a time, but several outgoing transition probabilities are reduced at once.

@@

We will phrase it in terms of $\snrb(\tau)$, reintroducing $r$:
%
\begin{equation}\label{eq:envunits}
  \snrb(\tau) = \frac{ \sqrt{N} \hat{A}\prn{\frac{1}{r\tau}} }{ r\tau },
  \qquad
  \alpha(\tau) = \frac{ 1 + r\tau +\sqrt{1+2r\tau} }{ r\tau }.
\end{equation}
%
Then, combining \eqref{eq:binaryenv}, \eqref{eq:unienv} and \eqref{eq:stickylaplace}, we find
%
\begin{equation}\label{eq:henv}
  \snrb(\tau) = \left\{
\begin{aligned}
  & \frac{\sqrt{N}}{1+r\tau}, &
  & r\tau \leq \frac{1}{2\sinh^2\frac{y_*}{2}} , \\
  & \sqrt{N}\prn{ \frac{ 4\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } } \log\alpha, &
  \frac{1}{2\sinh^2 \frac{y_*}{2}} \leq & r\tau \leq \frac{1}{2\sinh^2 \frac{y_*}{M}} , \\
  & \max_\varepsilon \brc{\frac{2\sqrt{N}\varepsilon}{2+(M-2)\varepsilon} \,
      \frac{ (\alpha^{\frac{M}{2}}-1)^2 - (1-\varepsilon)\alpha(\alpha^{\frac{M-2}{2}}-1)^2 }{ (\alpha^M + 1) - (1-\varepsilon)\alpha(\alpha^{M-2}+1) }}, &
  \frac{1}{2\sinh^2 \frac{y_*}{M}} \leq & r\tau .
\end{aligned}
\right.
\end{equation}
%
For $M \gg 1$, and $r\tau$ much greater than the lower bound of each region, we can approximate this as
%
\begin{equation}\label{eq:henvapproc}
  \snrb(\tau) \approx \left\{
\begin{aligned}
  & \frac{\sqrt{N}}{1+r\tau}, &
  & r\tau \leq 0.73 , \\
  & \frac{ 0.54\sqrt{N}}{ \sqrt{r\tau} } , &
  0.73 \leq & r\tau \leq 0.22 M^2 , \\
  & \frac{ \sqrt{N} (M-1) }{ r\tau }, &\quad
  0.22 M^2 \leq &r\tau .
\end{aligned}
\right.
\end{equation}
%
This is plotted in \autoref{fig:heuristicenv}.
